\documentclass[11pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\usepackage{pdfpages} 
\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{ulem}

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{graphicx, fancyhdr}
\usepackage{amsmath, amsfonts}
\usepackage{color}
\usepackage{hyperref}

\newcommand{\blue}[1]{{\color{blue} #1}}

\setlength{\topmargin}{-.5 in} 
\setlength{\textheight}{9 in}
\setlength{\textwidth}{6.5 in} 
\setlength{\evensidemargin}{0 in}
\setlength{\oddsidemargin}{0 in} 
\setlength{\parindent}{0 in}
\newcommand{\ben}{\begin{enumerate}}
\newcommand{\een}{\end{enumerate}}


\lhead{STAT 305}
\chead{Homework \# 8} 
\rhead{Due Thursday, Nov. $8^{th}$ in the class}
\lfoot{Fall 2019} 
\cfoot{\thepage} 
\rfoot{} 
\renewcommand{\headrulewidth}{0.4pt} 
\renewcommand{\footrulewidth}{0.4pt} 

\def\Exp#1#2{\ensuremath{#1\times 10^{#2}}}
\def\Case#1#2#3#4{\left\{ \begin{tabular}{cc} #1 & #2 \phantom
{\Big|} \\ #3 & #4 \phantom{\Big|} \end{tabular} \right.}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\usepackage{Sweave}
\begin{document}
\input{stat305_hw8-concordance}
\pagestyle{fancy} 

Show \textbf{all} of your work on this assignment and answer each question fully in the given context. \\


\emph{Please} staple your assignment!


\begin{enumerate}
	
    \item \textbf{Ch. 5, Exercise 7, pg. 323:} In a grinding operation, there is an upper specification of $3.150$ in. on a dimensions of a certain part after grinding. Suppose that the standard deviation of this normally distributed dimension for parts of this type ground to any particular mean dimension $\mu$ is $\sigma = 0.002$ in. Suppose further that you desire to have no more than $3\%$ of the parts fail to meet specifications. What is the maximum (minimum machining cost) $\mu$ that can be used if this $3$\% requirement is to be met?[10 pts]\\
    Hint: the question is giving you information on $P(X> 3.15)\leq 0.03$.
    
    \item \textbf{Ch 5, Exercise 42, pg. 332:} Suppose that engineering specifications on the shelf depth of a certain slug to be turned on a CNC lathe are from $0.0275$ in. to $0.0278$ in. and that values of this dimension produced on the lathe can be described using a normal distribution with mean $\mu$ and standard deviation $\sigma$.
        \begin{enumerate}
              \item If $\mu = 0.0276$ and $\sigma = 0.0001$, about what fraction of shelf depths are in specifications?[10 pts]
              
              \item What machine precision (as measured by $\sigma$) would be required in order to produce about $98$\% of shelf depths within engineering specifications (assuming that $\mu$ is at the midpoint of the specifications)?[10 pts]\\
              Hint: you are looking for the value of $\sigma$ in this question.
        \end{enumerate}
 \item \textbf{Ch 5.4, Exercise 2, pg. 300:} Quality audit records are kept on numbers of major and minor failures of circuit packs during burn-in of large electronic switching devices. They indicate that for a device of this type, the random variables
    $$
    X = \text{ the number of major failures}
    $$
    and
    $$
    Y = \text{ the number of minor failures}
    $$
    can be described at least approximately by the accompanying joint distribution.



 \begin{table}[h!]
     \centering
     \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|}
     \hline
         \textbf{Y | X}& \textbf{0} & \textbf{1}  & \textbf{2}  \\\hline \hline
     0 & 0.15  & 0.05  & 0.01  \\
     \hline
     1 & 0.1   & 0.08  &0.01  \\
     \hline
     2 & 0.1   & 0.14  & 0.02  \\
     \hline
     3 & 0.1   & 0.08  & 0.03  \\
     \hline
     4 & 0.05  & 0.05  & 0.03  \\
     \hline
     \end{tabular}
  \end{table}
          \begin{enumerate}
            \item Find the marginal probability functions for both $X$ and $Y$ ($f_x(x)$ and $f_y(y)$, respectively).[10 pts]
            \item Are $X$ and $Y$ independent? Explain.[5pts]
            \item Find the mean and variance of $X$ ($\text{E}X$ and $\text{Var}X$)[10 pts]
            \item Find the mean and variance of $Y$ ($\text{E}Y$ and $\text{Var}Y$)[10 pts]
            \item Find the conditional probability function for $Y$, given that $X = 0$ -- i.e., that there are no major circuit pack failures ($f_{Y|X}(y|0)$). What is the mean of this conditional distribution?[10 pts]
            \end{enumerate}
  
  \item \textbf{Ch. 5.2, Exercise 3, pg. 263:} Suppose that $X$ is a normal random variable with mean $43$ and standard deviation $3.6$. Evaluate the following probabilities involving $X$:
      \begin{enumerate}
              \item $P[X < 45.2]$ [5 pts]
              \item $P[|X - 43| \le 2]$ [5 pts]
              \item $P[|X - 43| > 1.7]$  [5 pts]
      
    Now find numbers \# such that the following statements involving $X$ are true:
    
              \item $P[X < \#] = .95$ [5 pts]
              \item $P[|X - 43| > \#] = .05$ [5 pts]

  
\end{enumerate}
Total: 90 pts



\end{document}
