\documentclass[11pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{ulem}

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{graphicx, fancyhdr}
\usepackage{amsmath, amsfonts}
\usepackage{color}
\usepackage{hyperref}

\newcommand{\blue}[1]{{\color{blue} #1}}

\setlength{\topmargin}{-.5 in} 
\setlength{\textheight}{9 in}
\setlength{\textwidth}{6.5 in} 
\setlength{\evensidemargin}{0 in}
\setlength{\oddsidemargin}{0 in} 
\setlength{\parindent}{0 in}
\newcommand{\ben}{\begin{enumerate}}
\newcommand{\een}{\end{enumerate}}


\lhead{STAT 305}
\chead{Homework 1- Solution} 
\rhead{}
\lfoot{Fall 2019} 
\cfoot{\thepage} 
\rfoot{} 
\renewcommand{\headrulewidth}{0.4pt} 
\renewcommand{\footrulewidth}{0.4pt} 

\def\Exp#1#2{\ensuremath{#1\times 10^{#2}}}
\def\Case#1#2#3#4{\left\{ \begin{tabular}{cc} #1 & #2 \phantom
{\Big|} \\ #3 & #4 \phantom{\Big|} \end{tabular} \right.}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\pagestyle{fancy} 
\begin{itemize}

\item \textbf{Problem1: Chapter 1, Exercise 2 (page 13)} \\
There are different possible answers, but one of them is:\\
Qualitative data: rating the quality of batches of ice cream as either poor, fair, good, or exceptional.\\
Quantitative data: measuring the time (in hours) it takes for each of 1000 integrated circuit chips to fail in a high-stress environment.


\item \textbf{Problem2: Chapter 1, Exercise 3 (page 13)}\\
Any relationship between the variables x and y can only be derived from a bivariate sample. \\

\item \textbf{Problem3: Chapter 1, Exercise 8 (page 24)}\\
(a) Rockwell hardness: multivariate (bivariate), repeated measures (paired), quantitative data. Flatness: univariate, qualitative data.\\
\\
(b) There are many possibilities. Possible factors are Vendor, Material, Heating time, Heating Temperature, Cooling Method, and Furance Atmosphere Condition. You could choose any number of levels for each factor. If you choose Vendor (1 VS. 2), Heating Time (short VS. Long), and Cooling Method (1 VS. 2), the factor-level combinations are given below:
\hspace{1in}
\begin{center}
	\begin{tabular}{|c|c|c|}\hline
		 Vendor & Heating Time & Cooling Method   \\ [0.5ex] 
		 \hline\hline
		1 & short&1 \\
		2 & short&1 \\
		1 & long&1  \\
		2 & long&1 \\
		1 & short&2 \\
		2 & short&2  \\
		1 & long&2  \\
		2 & long&2  \\
		\hline
	\end{tabular}
\end{center}
\hspace{1 in}


\item \textbf{Problem4: Chapter 2, Section 2, Exercise 3 (page 37)}\\
A simple random sample is not guaranteed to be representative of the population from which is drawn. It gives every set of n items an equal chance of being selected, so there is always a chance that the n items chosen will be "extreme" member s of the population. That is the discussion of biassness we had in the class. 

\item \textbf{Problem5: Hockey game attendance.}

Caroline performs the following study to see if outside temperature
has an effect on attendance at her college's hockey games. For each
hockey game at her college, Caroline records the outside temperature and the attendance. Here are her results:

\hspace{1in}
\begin{tabular}{|cr@{/}r|c|c|} \hline
\multicolumn{3}{|c|}{\emph{Day of Week}} & \emph{Temperature, deg. F} & \emph{Attendance} \\ \hline
Friday & 12&14 & 35 & 840 \\
Wednesday & 12&19 & 20 & 560 \\
Tuesday & 1&8 & $\!-5$ & 340 \\
Friday & 1&11 & 23 & 775 \\
Wednesday & 1&23 & 14 & 680 \\
Saturday & 2&2 & 30 & 950 \\
Friday & 2&8 & 28 & 950 \\
\hline
\end{tabular}
 
\ben 
\item Is this an experiment or observational study?\\
Observational study, the investigator (Caroline) has a passive role in the study. The temperature and the attendance are  watched and data
are recorded, but there is no intervention on the part of the person conducting the study
\item What type of variable is attendance?\\
Discrete quantitative (numerical)
\een

Caroline analyzes her results and finds that
outside temperature and attendance have a strong
positive correlation (i.e., as one increases, the other also increases).
She concludes that higher game day temperatures causes higher
attendance at their college's hockey games.

\ben
\addtocounter{enumii}{2}
\item Did she come to a proper conclusion for this study? Why or why not?\\
(Multiple answers are possible, if they are to the point, you get full credit for them)\\
In the first place it seems the answer is yes, but there are some hidden assumptions that might affect the study. Among many possible ones we can mention the following:\\
\begin{itemize}
\item The match day (weekend or weekdays)
\item This game might not be much popular among the people in that city.
\item The team migh not have done well in that season.
\item The match dates.
\item The importance of the match.
\item The opponent. 
\end{itemize}


\item Look at the day of the week of the hockey games.
What type of variable is this?\\
Qualitative (categorical)
 
\item 
Rewrite the data table, adding a new column ``School Night" (using the values ``no" if the game is on a Friday or Saturday, and ``yes" if the game is on any other day).
How does Attendance relate to School Night?
\hspace{1in}

\begin{tabular}{|cr@{/}r|c|c|c|} \hline
\multicolumn{3}{|c|}{\emph{Day of Week}} & \emph{Temperature, deg. F} & \emph{Attendance} & \emph{School Night}\\ \hline
Friday & 12&14 & 35 & 840 & No\\
Wednesday & 12&19 & 20 & 560& Yes \\
Tuesday & 1&8 & $\!-5$ & 340 & Yes\\
Friday & 1&11 & 23 & 775 & No\\
Wednesday & 1&23 & 14 & 680& Yes \\
Saturday & 2&2 & 30 & 950& No \\
Friday & 2&8 & 28 & 950 &No\\
\hline
\end{tabular}
\een

More people watch the game when the game is not on school nights. 
\item \textbf{Problem6: Washer stretching.}
  
George works for a company that manufactures 
rubber washers.  He randomly selects 1000 
washers off the assembly line throughout two weeks
for a study on the durability of these washers under  
stretching.  To make sure that the washers 
are fit to be used in the real world, George must test 
the washers.  Holding heat constant, 
George subjects each washer to one of various methods of 
stretching.  The washers are randomly assigned to be stretched under one of five different forces (low, medium-low, medium, medium-high, and high).
After each test, George classifies 
a washer as either defective or non-defective. 
 
\ben 
\item Is this an experiment or observational study? \\
Experimental study
\item What type of variable is heat? \\
Quantitative (numerical)
 
\item What type of variable is the amount of stretching? \\
Qualitative (categorical)
 
\item What type of variable is response to the stretching method?\\
Qualitative  (categorical)
 
\item
The 1000 selected washers constitutes the sample. What is the population? 
All rubber washers manufactured in the company

\item George analyzes the results and finds that the defect rate
increases with the amount of stretching.
Can George conclude that the amount of stretching 
causes a change in the defect rate of the washers?
Why or why not?\\


Remember, though, that there always exists the chance that the sample
was not properly representative of the population, and perhaps 
the 1000 washers he selected 
were some of the few washers in the population
that showed no effect. 
But through proper
experimental design, this chance is minimized and George can 
be highly confident in the results (this is again the matter of highly biassed discussion in the class, but this study was different with the example of lightbulbs as the sampling was randomly over two weeks). More on confidence in Chapter 6.
\een


\end{itemize}


\end{document}
