---
title: "Homework 8 (Practice)"
date: "Not due"
output: 
  pdf_document:
    number_sections: true
geometry: margin=1in
---

```{r setup, echo=FALSE, message=FALSE}
library(knitr)
library(tidyverse)
library(xtable)

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, fig.height = 2)
theme_set(theme_bw(base_family = "serif"))
```

1. [Ch. 9.1, Exercise 1, pg. 674] Return to the data from Homework 4, Exercise 1. The article "Polyglycol Modified Poly (Ethylene EtherCarbonate) Polyols by Molecular Weight Advancement" by R. Harris (*Journal of Applied Polymer Science*, 1990) contains some data on the effect of reaction temperature on the molecular weight of resulting poly polyols. The data for eight experimental runs at temperature 165°C and above are as follows (see website for `polyols.csv`):

    ```{r}
    polyols <- read_csv("../data/polyols.csv")
    
    kable(polyols, col.names = c("Pot temperature (°C)", "Average molecular weight"))
    ```
    
    a) Find $s_{LF}$ for these data. What does this intend to measure in the context of the engineering problem?
    c) Give a 90% two-sided confidence interval for the increase in mean average molecular weight that accompanies a 1°C increase in temperature here.
    d) Give individual two-sided confidence intervals for the mean average molecular weight at 212°C and also at 250°C.
    h) Use a hypothesis test (via p-value) with $\alpha = .05$ to determine if there is a significant linear relationship between molecular weight and temperature of poly polyols.
    
2. [Ch 9.2, Exercise 2, Pg. 697] Return to the study in Homework 5, Exercise 2. Here are some data (also on website as `pulp.csv`) taken from the article "Chemithermomechanical Pulp from Mixed High Density Hardwoods" by Miller, Shankar, and Peterson (*Tappi Journal*, 1988). Given are the percent NaOH used as a pretreatment chemical, $x_1$, the pretreatment time in minutes, $x_2$, and the resulting value of a specific surface area variable, $y$ (with units of $\text{cm}^2$/g), for nine batches of pulp produced from a mixture of hardwoods at a treatment temperature of $75$°C in mechanical pulping. 

    ```{r}
    pulp <- read_csv("../data/pulp.csv")
    
    kable(pulp, col.names = c("% NaOH, $x_1$", "Time, $x_2$", "Specific Surface Area, $y$"))
    ```
    
    Consider an analysis of the data based on the model $y = \beta_0 + \beta_1x_1 + \beta_2x_2 + \epsilon$.
    
    a) Find $s_{SF}$. What does this intend to measure in the context of the engineering problem?
    c) Give 90% individual two-sided confidence intervals for all of $\beta_0$, $\beta_1$, and $\beta_2$.
    d) Give individual 90% individual two-sided confidence intervals for the mean specific surface area, first when $x_1 = 9.0$ and $x_2 = 60$ and then when $x_1 = 10.0$ and $x_2 = 70$.
    h) Use a hypothesis test (via p-value) with $\alpha = .05$ to determine if there is a significant linear relationship between specific surface area weight and time, with % NaOH held fixed.